A Reproducible Data Analysis Workflow with RMarkdown, Git, Make & Docker

### 2020-05-13 manuscript.Rmd
revision of Make example to increase clarity

A `Makefile` obeys a simple scheme. It contains a list of recipes. Each recipe has a target (the name of the recipe) followed by a list of dependent targets or files. If any of the dependencies have changed since the last time the target was built, the recipe's commands are executed to recreate the target file. A typical `Makefile` might look like this:

```{bash, eval=FALSE}
analysis.pdf: data/clean.csv
  Rscript -e 'rmarkdown::render("analysis.Rmd")'

data/clean.csv: R/clean.R data/iris.csv
  Rscript -e 'source("R/clean.R")'
```

The first line can be interpreted as follows: To create `analysis.pdf` (the first target in the file), the file `analysis.Rmd`, which depends on `data/clean.csv`, needs to be rendered. This dependency is itself a target (the second target in the file). To create `data/clean.csv`, `R/clean.R` and `data/iris.csv` are needed. If you type `make analysis.pdf`, Make first checks whether the dependencies do exist and, if not, creates them. Here, if `data/clean.csv` does not exist, Make creates it by executing the second target. Also, if one of the dependencies of a target is newer then the target itself, then Make updates everything that directly or indirectly depends on the target. Here, if `data\iris.csv` is newer then `data\clean.csv`, Make will attempt to recreate `data\clean.csv` first and then `analysis.pdf`. If there is a dependency which is missing, and there is no target to make it, Make stops with an error message. It is a convention to have the first target named `all`, which depends on everything. Then the command `make` without any argument automatically creates everything possible in the project. The button `Build All` from within RStudio triggers this process (see Figure \@ref{fig:buildpane-rstudio}).

>>>>

A `Makefile` obeys a simple scheme. It contains a list of recipes. Each recipe has a target (the name of the recipe) followed by a list of dependent targets or files. If any of the dependencies have changed since the last time the target was built, the recipe's commands are executed to recreate the target file. We illustrate the use of Makefiles with an example. Assume the final product is a report (`report.pdf`). This report is written in Rmarkdown (`report.Rmd`) and includes dynamically generated plots from a raw data file (`data/iris.csv`) that needs to be preprocessed first using a separate script (`R/preprocess.R`) into a processed data file (`ìris-processed.csv`). A `Makefile` for these dependencies may look like this:

```{bash, eval=FALSE}
report.pdf: data/iris-processed.csv report.Rmd
  Rscript -e 'rmarkdown::render("report.Rmd")'

data/iris-processed.csv: R/preprocess.R data/iris.csv
  Rscript -e 'source("R/preprocess.R")'
```

The first line can be interpreted as follows: To create `report.pdf` (the first target in the file), the file `report.Rmd`, which depends on `data/iris-processed.csv`, needs to be rendered. This dependency is itself a target (the second target in the file). To create `data/iris-processed.csv`, `R/preprocess.R` and `data/iris.csv` are needed. If you type `make report.pdf`, Make first checks whether the dependencies do exist and, if not, creates them. Here, if `data/iris-processed.csv` does not exist, Make creates it by executing the second target. Also, if one of the dependencies of a target is newer then the target itself, then Make updates everything that directly or indirectly depends on the target. Here, if `data\iris.csv` is newer then `data\iris-processed.csv`, Make will attempt to recreate `data\iris-processed.csv` first and then `report.pdf`. If there is a dependency missing, and there is no target to make it, Make stops with an error message. It is a convention to have the first target named `all`, which depends on everything. Then the command `make` without any argument automatically creates everything possible in the project. The button `Build All` from within RStudio triggers this process (see Figure \@ref{fig:buildpane-rstudio}).
